---
title: "DNAm and Wildfire PM2.5, Enrichment analyses"
author: "Rongbin Xu"
date: "27 Oct 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,message=FALSE}
rm(list = ls())
library(dplyr)
library(missMethyl)
```

## Read EWAS and DMR results
```{r}
## EWAS results
EWAS_compare<-readRDS("Wildfire_EWAS_reuslts.rds")  
exp<-levels(EWAS_compare$exposure)
sig_CpGs_Bonferi<-EWAS_compare%>%
  filter(P_bonferroni<0.05)

## DMR results
over_DMRcate_0.01<-read.csv("DMRcate_overlaps_p0.01_order_by_location.csv")[,-1]
over_comb_p_0.01<-read.csv("Comb_p_overlaps_p0.01_order_by_location.csv")[,-1]
```

## GO and KEEG by missMethyl
```{r}
## Prepare data for missMethyl
library(IlluminaHumanMethylation450kanno.ilmn12.hg19)
library(limma)
ann <- getAnnotation(IlluminaHumanMethylation450kanno.ilmn12.hg19)

```

### Eenrichment analyses
Here, I chose inputs as CpGs(Bonferroni P-value<0.05) from EWAS and CpGs overlaped with DMRs by two methods (FDR<0.01).
```{r}
CPG_backgroud<-unique(EWAS_compare$probeID)

## Analyses 1
input1_wildfire<-c(as.character(sig_CpGs_Bonferi$probeID[sig_CpGs_Bonferi$exposure=="Wildfire-related"]),
          as.character(over_comb_p_0.01$CpG[over_comb_p_0.01$exposure=="Wildfire-related"]),
          as.character(over_DMRcate_0.01$CpG[over_DMRcate_0.01$exposure=="Wildfire-related"]))%>%
          paste(collapse = "/")%>%
          strsplit("/")%>%
          unlist()%>%
          unique()

input1_non_wildfire<-c(as.character(sig_CpGs_Bonferi$probeID[sig_CpGs_Bonferi$exposure=="Non-wildfire-related"]),
          as.character(over_comb_p_0.01$CpG[over_comb_p_0.01$exposure=="Non-wildfire-related"]),
          as.character(over_DMRcate_0.01$CpG[over_DMRcate_0.01$exposure=="Non-wildfire-related"]))%>%
          paste(collapse = "/")%>%
          strsplit("/")%>%
          unlist()%>%
          unique()



output1_GO_wildfire<-gometh(input1_wildfire,collection = "Go",anno = ann,all.cpg = CPG_backgroud,sig.genes = TRUE)%>%
                     mutate(exposure="Wildfire-related",ID=rownames(.))## about 10 mins
output1_GO_non_wildfire<-gometh(input1_non_wildfire,collection = "Go",anno = ann,all.cpg = CPG_backgroud,sig.genes = TRUE)%>%
                         mutate(exposure="Non-wildfire-related",ID=rownames(.))

output1_KEEG_wildfire<-gometh(input1_wildfire,collection = "KEGG",anno = ann,all.cpg = CPG_backgroud,sig.genes = TRUE)%>%
                       mutate(exposure="Wildfire-related",KEGG_id=rownames(.))
output1_KEEG_non_wildfire<-gometh(input1_non_wildfire,collection = "KEGG",anno = ann,all.cpg = CPG_backgroud,sig.genes = TRUE)%>%
                           mutate(exposure="Non-wildfire-related",KEGG_id=rownames(.))




GO_enrich<-output1_GO_wildfire%>%
  bind_rows(output1_GO_non_wildfire)
saveRDS(GO_enrich,file="GO_enrichment_results.rds")


KEEG_enrich<-output1_KEEG_wildfire%>%
  bind_rows(output1_KEEG_non_wildfire)
saveRDS(KEEG_enrich,file="KEEG_enrichment_results.rds")
```



## Process enrichment results
```{r}
GO_enrich<-readRDS("GO_enrichment_results.rds")
KEEG_enrich<-readRDS("KEEG_enrichment_results.rds")

sig_GO<-GO_enrich%>%
  filter(DE>=3&P.DE<0.05)%>%#### significantly enriched pathway: al least three genes, and P-value for enrichment<0.05
  group_by(exposure,SigGenesInSet)%>%
  arrange(exposure,P.DE)
xtabs(~exposure+ONTOLOGY,sig_GO)

topGSA(filter(sig_GO,exposure=="Wildfire-related"),number = 10) 
topGSA(filter(sig_GO,exposure=="Non-wildfire-related"),number = 10) 
shared_GO<-intersect(filter(sig_GO,exposure=="Wildfire-related")$TERM,
                          filter(sig_GO,exposure=="Non-wildfire-related")$TERM)## none





sig_KEEG<-KEEG_enrich%>%
  filter(DE>=3&P.DE<0.05)%>%#### significantly enriched pathway: al least three genes, and P-value for enrichment<0.05
  group_by(exposure,SigGenesInSet)%>%
  arrange(exposure,P.DE)
topGSA(filter(sig_KEEG,exposure=="Wildfire-related"),number = 10) 
topGSA(filter(sig_KEEG,exposure=="Non-wildfire-related"),number = 10) 
  
shared_KEEG<-intersect(filter(sig_KEEG,exposure=="Wildfire-related")$TERM,
                          filter(sig_KEEG,exposure=="Non-wildfire-related")$TERM)## none

```
## Adapt simplify fuction from  Clusterprofiler package
It is clear from above results that many GO terms actually contains the same gene set. This is becasue many GO terms have complicated parent-child relationships. Reporting all these GO terms would hinder effective intepretation. The gometh function in missMethyl has not address this problem. 

However, the "clusterProfiler" R package provides a function called "simplify" to eliminate redundant GO terms (see Wu et al. The Innovation 2021;2:100141). This function employs GOSemSim package to calculate semantic similarities (an indicator betwen 0 and 1) among enriched GO terms using multiple methods based on information content or graph structure. For each group of GO terms that have high similarity between each other (similarity >0.7), the simplify function would only retain a repsentative term (e.g., the most significantly enriched term) while remove the others.

see https://github.com/YuLab-SMU/clusterProfiler/commit/7e32c7d180ca928a22b58755d2949ad429c1e45f for the original "simplify" function
```{r}
library(clusterProfiler)
library(GOSemSim)
library(dplyr)
x<-sig_GO%>%
  mutate(exposure="Wildfire-related")%>%
  filter(ONTOLOGY=="BP")## try one ontology as an example, because the original simplify function performs similarity calculation by ontology.

## function for processing one specific ontology  
simplify_one_on<-function(res=x, cutoff=0.7, by="P.DE", select_fun=min, measure="Wang",computeIC=FALSE) {
              ## to satisfy codetools for calling gather
              go1 <- go2 <- similarity <- NULL

              sim <- mgoSim(res$ID, res$ID,
                            semData = godata('org.Hs.eg.db', ont=unique(res$ONTOLOGY), computeIC=computeIC),
                            measure=measure,
                            combine=NULL)

              sim.df <- as.data.frame(sim)
              sim.df$go1 <- row.names(sim.df)
              sim.df <- tidyr::gather(sim.df, go2, similarity, -go1)

              sim.df <- sim.df[!is.na(sim.df$similarity),]

              ## feature 'by' is attached to 'go1'
              sim.df <- merge(sim.df, res[, c("ID", by)], by.x="go1", by.y="ID")
              sim.df$go2 <- as.character(sim.df$go2)

              ID <- res$ID

              GO_to_remove <- character()
  
              for (i in seq_along(ID)) {
                  ii <- which(sim.df$go2 == ID[i] & sim.df$similarity > cutoff)
                  ## if length(ii) == 1, then go1 == go2
                  if (length(ii) < 2) 
                      next

                  sim_subset <- sim.df[ii,]

                  jj <- which(sim_subset[, by] == select_fun(sim_subset[, by]))
                  
                  if(length(jj)>1){## Some GO terms may have exact the same P-value, in this case, select the GO term with longer names
                   s_jj<-which(nchar(res[res$ID%in%sim_subset$go1[jj],]$TERM)==max(nchar(res[res$ID%in%sim_subset$go1[jj],]$TERM)))
                   jj<-jj[s_jj]
                  }

                  ## sim.df <- sim.df[-ii[-jj]]
                  GO_to_remove <- c(GO_to_remove, sim_subset$go1[-jj]) %>% unique
              }
           
              filter(res,!ID%in%GO_to_remove)
              
}



x<-simplify_one_on(x,measure="Jiang",computeIC = T)
## after try severl meaures, the "Jiang" looks like the best to remove redudant terms
## "Wang" measure doen't work for BP terms(all similarity equal to 0)
```
## Simplfied GO enrichment results
```{r}
simple_GO_wildfire<-sig_GO%>%
  filter(exposure=="Wildfire-related")%>%
  split.data.frame(.$ONTOLOGY)%>%
  lapply(function(a) simplify_one_on(a,cutoff=0.7, by="P.DE", select_fun=min, measure="Jiang",computeIC = T))%>%
  bind_rows()

simple_GO_non_wildfire<-sig_GO%>%
  filter(exposure=="Non-wildfire-related")%>%
  split.data.frame(.$ONTOLOGY)%>%
  lapply(function(a) simplify_one_on(a,cutoff=0.7, by="P.DE", select_fun=min, measure="Jiang",computeIC = T))%>%
  bind_rows()
## Save results
simple_GO_wildfire%>%
  bind_rows(simple_GO_non_wildfire)%>%
write.csv("GO_singnificant_enriched_pathways.csv")

sig_KEEG%>%
  write.csv("KEEG_singnificant_enriched_pathways.csv")


```

## try plot significant pathways
```{r}
library(ggplot2)
library(ggsci)
A<-simple_GO_wildfire%>%
  mutate(y=-log10(P.DE))%>%
  arrange(ONTOLOGY,y)%>%
  mutate(TERM=factor(TERM,levels = .$TERM))%>%
  ggplot(aes(x=TERM,y=y,fill=ONTOLOGY))+
  scale_fill_lancet()+
    geom_bar(stat="identity")+
    coord_flip()+
    labs(x="",y=expression(paste("-log"[10],"(P-value)",sep = "")),fill="GO ontology",
         title = "A.Wildfire-related")+
    theme_bw()
A
```

```{r}
B<-simple_GO_non_wildfire%>%
  mutate(y=-log10(P.DE))%>%
  arrange(ONTOLOGY,y)%>%
  mutate(TERM=factor(TERM,levels = .$TERM))%>%
  ggplot(aes(x=TERM,y=y,fill=ONTOLOGY))+
  scale_fill_lancet()+
    geom_bar(stat="identity")+
    coord_flip()+
    labs(x="",y=expression(paste("-log"[10],"(P-value)",sep = "")),fill="GO ontology",
         title = "B.Non-wildfire-related")+
    theme_bw()
B

## seem that it is better to display the results as table, rather as figures
```


